---
title: "Creation of an INDEF query file"
output: html_notebook
---

# Introduction

This notebook creates a sample file to query the INDEF registry.
It follows the instructions in [INDEF's User Manual][UM].
More specifically, in section 7, modality "collective query" (p. 6).

[UM]: https://www.mscbs.gob.es/estadEstudios/estadisticas/estadisticas/estMinisterio/docs/Manual_INDEF_2017.pdf


# Composing the file

The manual says that the fields must be separated by the pipe `|` character.
There are options to search by 1, 4, 7 or 8 fields.
Within the same file, fields may be empty for a row,
just by putting two pipes one after another.
Therefore, it seems reasonable to use the 8-field format,
which admits the other ones and is thus the most flexible one.

In the following code I try to create a file
using the 8-field format with bogus data.

```{r}
if (!require(pacman)) install.packages("pacman")
library(pacman)
p_load(tidyverse, lubridate)


deaths_data <- tibble(
  name       = c("Daniel", "Obi-Wan"),
  surname_1  = c("Morillo", "Kenobi"),
  surname_2  = c("Cuadrado", NA),
  sex        = c('v', 'm'),
  birth_date = dmy(c('29-05-1981', '06-01-1977')),
  id_num     = c("50116894D", "X4661287F"),
  id_type    = c('D', 'N')
)

# Get necessary fields:

query_data <- deaths_data |>
  mutate(
    birth_day   = birth_date |> day(),
    birth_month = birth_date |> month(),
    birth_year  = birth_date |> year(),
    final       = NA_character_ # This forces an additional pipe at the end
  )                       |>
  select(name:sex, birth_day, birth_month, birth_year, id_num, id_type, final)
```


`write_delim()` is the canonical function to write delimited text files in
the _tidyverse_; however, it does not admit an encoding parameter,
and output files are always encoded in `UTF-8` according to its help page.
Therefore, we use `utils::write.table()` instead

```{r}
query_data |> write.table(
  file         = "../dat/indef_query_data.dat",
  quote        = FALSE,
  sep          = '|',
  na           = '',
  row.names    = FALSE,
  col.names    = FALSE,
  fileEncoding = 'Windows-1252'
  # (Stands for 'ANSI' encoding; see https://stackoverflow.com/a/701920/1585338)
)
```

